---
title: "Homework 3"
author: Minjie Bao
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup,include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() +  theme(legend.position = "bottom"))

options(
  ggplots2.continuous.color = "viridis",
  ggplots2.continuous.fill = "viridus"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



### Problem 2
```{r part1}
accel_df =
read_csv("./data/accel_data.csv") %>% 
pivot_longer(
  activity.1 : activity.1440,
    names_to = "activity",
    values_to = "minutes"
) %>% 
mutate(
  weekday=case_when(
day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
day %in% c("Saturday","Sunday") ~ "weekend"
),
day = factor(day),
day = forcats::fct_relevel(day, c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
weekday = as.factor(weekday) 
)

head(accel_df)
count(accel_df)
```

There are 6 variables in the cleaned dataset: `r names(accel_df)`. This dataset contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. There are `r count(accel_df)` observations. The variables "week", "day_id" and "minutes"are double datat ype. "day" and "weekday" are factor datat ype. "activity" is character data ype.


```{r part2}
accel_df %>% 
group_by(week, day, weekday) %>% 
  summarize(total_activity = sum(minutes)) %>% 
    knitr::kable(digits = 1)
```

For week 1, the 63 year-old male has the least activity on Monday, and  has the most activity on Sunday.

For week 2, the 63 year-old male has the least activity on Monday, and  has the most activity on Saturday.

For week 3, the 63 year-old male has the least activity on Thursday, and  has the most activity on Monday.

For week 4, the 63 year-old male has the least activity on Saturday, and  has the most activity on Wednesday.

For week 5, the 63 year-old male has the least activity on Saturday, and  has the most activity on Friday.

### Problem 3


