---
title: "Homework 3"
author: Minjie Bao
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup,include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() +  theme(legend.position = "bottom"))

options(
  ggplots2.continuous.color = "viridis",
  ggplots2.continuous.fill = "viridus"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user/ order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.

How many aisles, and which are most items from?
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

 Let's make a plot
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n>10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle,n)
  ) %>% 
  ggplot(aes(x=aisle, y=n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
 
Let's make a table!!
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Apples vs ice cream..
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```



### Problem 2
```{r p2_part1}
accel_df =
read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
pivot_longer(
  activity_1 : activity_1440,
    names_to = "minute",
    values_to = "activity_count"
) %>% 
mutate(
  weekday=case_when(
day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
day %in% c("Saturday","Sunday") ~ "weekend"
),
day = factor(day),
day = forcats::fct_relevel(day, c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
weekday = factor(weekday) 
)


head(accel_df)
count(accel_df)
```

There are 6 variables in the cleaned dataset: `r names(accel_df)`. This dataset contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. There are `r count(accel_df)` observations. The variables: "week", "day_id" and "activity_count" are double data type. "day" and "weekday" are factor data type. "minute" is character data type.


```{r p2_part2}
accel_df %>% 
group_by(week, day) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
    knitr::kable(digits = 2)
```

There's no apparent trends, but it seems like Tuesday and Wednesday's total activity is more stable than other days.

```{r p2_part3}
accel_df %>% 
    separate(minute, 
           into = c("char", "minute"),
           convert = TRUE
           ) %>% 
  select (-c(char)) %>% 
ggplot(aes(x=minute, y=activity_count,color = day)) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(title = "24-hour activity time")
```

From the plot we can see that this 63 year-old male has more activity on Friday night and Sunday morning than other days.

### Problem 3
```{r p3}
data("ny_noaa") 
```

This data set contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The size of the data set is `r nrow(ny_noaa)` * `r ncol(ny_noaa)`. There are 7 variables: `r names(ny_noaa)`. There' are's an issue that too many missing values in tmax and tmin. Variable "tmin" has `r sum(is.na(pull(ny_noaa, tmin)))`missing values and variable "tmax" has `r sum(is.na(pull(ny_noaa, tmax)))` missing values.

```{r p3_part1}
ny_noaa = ny_noaa %>% 
separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10
  ) 
print(ny_noaa)

snow_count = count(ny_noaa, snow) %>% 
  mutate(
    snowfall_rank = min_rank(desc(n))
    )%>% 
arrange(snowfall_rank, by_group = FALSE)
print(snow_count)
```

In this step, I covert variables: tmax and tmin from character to numeric because these two variables are temperature. I also divided prcp, tmax and tmin by 10 to make them in resonable units. Now prcp has unit: mm, tmax and tmin has units: degrees C.

For snowfall, the most commonly observed values is 0 mm because most of the timefrom January 1, 1981 through December 31, 2010, New York state doesn't snow.The second commonly observed values is NA because there are `r sum(is.na(pull(ny_noaa, snow)))` missing values. The third commonly observed values is 25 mm because ny state normally doesn't rain a lot through the 30 years.

```{r p3_part2}
ny_noaa %>% 
  group_by(id, year, month) %>% 
  summarise(avg_tmax = mean(tmax, na.rm = TRUE)) %>% 
  filter(month %in% c(1,7)) %>% 
  drop_na() %>% 
ggplot(aes(x = year, y = avg_tmax)) +
geom_point() + 
geom_path(alpha = 0.1) +
  geom_smooth(se = TRUE) +
facet_grid(.~ month) + 
labs(title = " average max temperature in January and July in each station across years", x = "year", y = "average max temperature (C)") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

The average max temperature in July is higher than January. In January, we can see that the average max temperature in each station across the years is oscillating. In July, the average max temperature in each station across the years is more stable than January. Therefore, we cannot conclude that the temperature is warming across years or global warming. There are many outliers in the graph such as the points around January 1982, and July 1988.

```{r p3_part3}
tmax_tmin_plot =
ny_noaa %>% 
  ggplot(aes(x = tmax, y = tmin)) +
geom_hex() +
    labs(title = "tmax_vs_tmin")

snowfall_plot = 
ny_noaa %>% 
filter(snow>0 & snow<100) %>% 
  ggplot(aes(x = year, y = snow, group = year)) +
geom_boxplot() +
  labs(title = " snowfall_plot")

tmax_tmin_plot + snowfall_plot
```


